---
title: "MCD12_difftesting"
format: html
editor: visual
---

# MCD12s v 6.1 Difference Testing

Retrieve MCD12 Q1 and Q2 v.6.1 from both [opendap](https://opendap.cr.usgs.gov/opendap/hyrax/MCD12Q2.061/h16v07.ncml.html) and [data pool](https://e4ftl01.cr.usgs.gov/). Then compare values between files to make sure the ones coming out of data pool are the same as the ones on opendap.

## OPeNDAP NetCDF

Using data from Senegal (h16v07) in year 2015 (opendap time request 14:1:14, because start at zero instead of 1). Open and retrieve Land Cover Type1 and QA data from netCDF

```{r}
# Load required libraries
  library(raster) 
  library(ncdf4)

# Open netcdf file using ncdf4 package 
  nc <- ncdf4::nc_open('C:/Users/blind/TEMP_workspace/testing/opendap_MCD12/h16v07_Q1go.ncml.nc4')
  
# Describe contents of nc file
  attributes(nc$var)$names
        print(paste("The file has",nc$nvar, "variables"))
        print(paste("The file has",nc$ndims, "dimensions"))
        print(paste("The file has",nc$natts, "NetCDF attributes"))

# Retrieve data array from file
  LCType1 <- ncvar_get(nc, attributes(nc$var)$names[2])
  QC1 <- ncvar_get(nc, attributes(nc$var)$names[1])
```

Also retrieve data from MCD12Q2

```{r}
# Open netcdf file using ncdf4 package 
  nc2 <- ncdf4::nc_open('C:/Users/blind/TEMP_workspace/testing/opendap_MCD12/h16v07_Q2go.ncml.nc4')
  
# Describe contents of nc file
  attributes(nc2$var)$names
        print(paste("The file has",nc2$nvar, "variables"))
        print(paste("The file has",nc2$ndims, "dimensions"))
        print(paste("The file has",nc2$natts, "NetCDF attributes"))

# Retrieve data array from file
  Sen <- ncvar_get(nc2, attributes(nc2$var)$names[2])
  QC <- ncvar_get(nc2, attributes(nc2$var)$names[1])
```

## Data Pool

```{r}
# Can also use neytCDF package for hdf files 
  hd <- ncdf4::nc_open('C:/Users/blind/TEMP_workspace/testing/opendap_MCD12/MCD12Q1.A2015001.h16v07.061.2022166014721.hdf')
  
   attributes(hd$var)$names
   LCType1_hd <- ncvar_get(hd, attributes(hd$var)$names[1])
   QC1_hd <- ncvar_get(hd, attributes(hd$var)$names[12])
```

Also retrieve data from MCD12Q2

```{r}
# Can also use neytCDF package for hdf files 
  hd2 <- ncdf4::nc_open('C:/Users/blind/TEMP_workspace/testing/opendap_MCD12/MCD12Q2.A2015001.h16v07.061.2022111215125.hdf')
  
   attributes(hd2$var)$names
   Sen_hd <- ncvar_get(hd2, attributes(hd2$var)$names[6])
   QC_hd <- ncvar_get(hd2, attributes(hd2$var)$names[12])
```

## Compare values in the two Arrays

Can check if arrays are the same in two ways: 1) identical and 2) ==

```{r}
# using identical in base R
identical(LCType1, LCType1_hd)
identical(QC1, QC1_hd)

identical(Sen, Sen_hd)
identical(QC, QC_hd)

# using "==" in base R. If LC_Type1 equals LC_Type1_hd, put 0, else put 1.  If there are no differences between the arrays, the sum of should be 0; if the sum is greater than 0, there are some differences.
diffs <- ifelse(LCType1==LCType1_hd, 0, 1)
sum   <- sum(diffs) 
sum
```
